Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof

### 2014-02-05 paper/Introduction.tex

\section{Introduction}
\label{sec:Introduction}

@@

For this approach to be truly viable, designers cannot be handed a
single proof of correctnessof a single design; rather, they need a
proof framework covering any variation of a design that might be of
interest for performance reasons. This would enable implementation
level changes without compromising on the ability to verify
correctness of the design.

@@

\noindent\textbf{Paper Outline}. In Section~\ref{Sec:Background} we
discuss our notion of correctness for our cache coherence protocol
framework. We discuss our MSI protocol example and its implementation
in Section~\ref{Sec:System}. In Section~\ref{Sec:ProofStructure} we discuss
the structuring of our proof framework and verification of the base
protocol. We show implementation refinements in
Section~\ref{Sec:Refinements} and conclude with a general discussion
of the applicability of our technique.

>>>>

\section{Introduction}
\label{Sec:Introduction}

@@

For this approach to be truly viable, designers cannot be handed a
single proof of correctness of a single design; rather, they need a
proof framework covering any variation of a design that might be of
interest for performance reasons. This would enable implementation
level changes without compromising on the ability to verify
correctness of the design.

@@

\noindent\textbf{Paper Outline}. In Section~\ref{Sec:Background} we
discuss our notion of correctness for our cache coherence protocol
framework. We discuss our MSI protocol example and its implementation
in Section~\ref{Sec:System}. In Sections~\ref{Sec:Safety} and
\ref{Sec:Liveness} we discuss the structuring of our proof framework and
verification of the base protocol. We show implementation refinements
in Section~\ref{Sec:Refinements} and conclude with a general
discussion of the applicability of our technique.

### 2014-02-05 paper/System.tex
some modifications of backgrounda
Introduction put back in. fix some links

\section{MSI-based cache coherence protocol}
\label{sec:System}

@@

Figure~\ref{parentside} shows the similar sequence of transitions that results
in downgrading a cache's directory entry for one of its chilren $c$. The
difference is in the guard for ChildRecvReq transition: the directory entry for
each of the $c$'s children should not be greater than the requested
downgrade, in order for $c$ to accept a request and send a downgrade
response.

Figure~\ref{childextra} shows transitions that happen in a non-root cache
because of the presence of \emph{voluntary} responses. This models eviction of
a location from a cache to make room for a new location.  This eviction usually
causes the permission state of the cache for that address to go to $\In$, while
we allow any downgrade to take place voluntarily.  We use the same \cpResp{}
FIFO for such a message. Such a voluntary response for an address can take
place only if the cache is not waiting for a response from its parent for
that address (ChildVolResp).

Because of the presence of voluntary response, a child could have already
downgraded to a state required by a downgrade request from its parent. In such
a case, the child simply drops the downgrade request (ChildDropReq).

@@

\begin{subfigure}{\textwidth}
\centering
\begin{tabular}{|ll|}
\hline
\multicolumn{2}{|p{\textwidth}|}{\textbf{ChildVolResp}$(c, x, a)$: Child $c$
sending a response to downgrade to $x$ for address $a$ without any request from
its parent}\\
\hline
Guard: & $(\forall i, \parent(i, c) \rightarrow c.\dstate[i][a] \le x) \wedge x < c.\state[a] \wedge c.\wait[a] = \False$\\
\hline
Action: & $\enq(\cpResp[c], \langle c.\state[a], x, a,%$\\
%& $\;\;\;%
\myif c.\state[a] = \Mo \mythen c.\data[a] \myelse \_\rangle)$;\\
& $c.\state[a] \Leftarrow x$;\\
\hline
\hline
\multicolumn{2}{|p{\textwidth}|}{\textbf{ChildDropReq}$(c)$: Child $c$
receiving a request from its parent and has already downgraded}\\
\hline
Guard: & 
%$\avail(\pc[c]) \wedge \mylet q := \first(\pc[c]) \myin q.\typ = \Req \wedge%$\\
%& $%
%q.\myto \ge c.\state[q.\addr]$\\
$\avail(\pc[c]) \wedge \mylet q := \first(\pc[c]) \myin q.\typ = \Req \wedge q.\myto \ge c.\state[q.\addr]$\\
\hline
Action: & $\deq(\pc[c])$;\\
\hline
\end{tabular}
\caption{Other transitions at a non-root cache}
\label{childextra}
\end{subfigure}

>>>>

\section{MSI-based cache coherence protocol}
\label{Sec:System}

@@

Figure~\ref{parentside} shows the similar sequence of transitions that results
in downgrading a cache's directory entry for one of its children $c$. The
difference is in the guard for ChildRecvReq transition: the directory entry for
each of the $c$'s children should not be greater than the requested
downgrade, in order for $c$ to accept a request and send a downgrade
response.

Figure~\ref{childextra} shows transitions that happen in a non-root
cache due to the presence of \emph{voluntary} responses. This models
eviction from a cache. This eviction usually causes the permission
state of the cache for that address to go to $\In$, while we allow any
downgrade to take place voluntarily.  We use the same \cpResp{} FIFO
for such a message. Such a voluntary response for an address can take
place only if the cache is not waiting for a response from its parent
for that address (ChildVolResp).

Because of the presence of voluntary responses, a child may have already
downgraded to a state required by a downgrade request from its parent. In such
a case, the child drops the downgrade request (ChildDropReq).

@@

\begin{subfigure}{\textwidth}
\centering
\begin{tabular}{|ll|}
\hline
\multicolumn{2}{|p{\textwidth}|}{\textbf{ChildVolResp}$(c, x, a)$: Child $c$
sending a response to downgrade to $x$ for address $a$ without any request from
its parent}\\
\hline
Guard: & $(\forall i, \parent(i, c) \rightarrow c.\dstate[i][a] \le x) \wedge x < c.\state[a] \wedge c.\wait[a] = \False$\\
\hline
Action: & $\enq(\cpResp[c], \langle c.\state[a], x, a,%$\\
%& $\;\;\;%
\myif c.\state[a] = \Mo \mythen c.\data[a] \myelse \_\rangle)$;\\
& $c.\state[a] \Leftarrow x$;\\
\hline
\hline
\multicolumn{2}{|p{\textwidth}|}{\textbf{ChildDropReq}$(c)$: Child $c$
receiving a request from its parent and has already downgraded}\\
\hline
Guard: & 
$\avail(\pc[c]) \wedge \mylet q := \first(\pc[c]) \myin q.\typ = \Req \wedge q.\myto \ge c.\state[q.\addr]$\\
\hline
Action: & $\deq(\pc[c])$;\\
\hline
\end{tabular}
\caption{Other transitions at a non-root cache}
\label{childextra}
\end{subfigure}
